Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Background.tex
Spelling and grammar in Intro + System

Before we can discuss our protocol and proof framework, we must first establish
exactly what it means for a cache coherence protocol to be correct.

Intuitively, the simplest view of shared memory is as an array of addressible
objects in which a single entry can be read or written by one of the many
processors at a time.  In this view, all operations in the system appear to
have a single unique total ordering.

@@

\subsection{Formal Definition of Store Atomicity}
A memory subsystem can be modeled as the following function. Its input models a
set of infinite number of requests ever sent by any processor, and its output
models a set of (possibly infinite number of) responses ever sent by the memory
subsystem to any processor.

@@

Each output of the memory subsystem contains a set (called \Response) of
elements containing the fields (a) \labelR, containing a pair from the set
$Processor*\mathbb{N}$, which is meant to identify a request uniquely, (b)
\dataR, meant to model the data returned from the memory subsystem and (c)
\timeR, a value from set $\mathbb{N}$, meant to represent the ``time'' at which
the corresponding request (identified by \labelR) is handled by the memory
subsystem. \timeR{} does not have a counterpart in a real implementation, it is
present solely as a specificational device to capture store atomicity. Since a
request corresponding to a response is identified using the field \labelR, for
notational convenience, we will access the fields of the corresponding request
directly from the response. For instance, $r.\addrQ$ for $r \in \Response$
denotes the $(r.\labelR).\addrQ$, \etc.

>>>>

Before we can discuss our protocol and proof framework, we must first
establish exactly what it means for a cache coherence protocol to be
correct. Intuitively, the simplest view of shared memory is as an
array of addressable objects in which a single entry can be read or
written by one of the many processors at a time.  In this view, all
operations in the system appear to have a single unique total
ordering.

@@

\subsection{Formal Definition of Store Atomicity}

A memory subsystem can be modeled as the following function. Its input models a
set of infinite number of requests ever sent by any processor, and its output
models a set of (possibly infinite number of) responses ever sent by the memory
subsystem to any processor.

@@

Each output of the memory subsystem contains a set (called \Response) of
elements containing the fields (a) \labelR, containing a pair from the set
$Processor*\mathbb{N}$, which is meant to identify a request uniquely, (b)
\dataR, meant to model the data returned from the memory subsystem and (c)
\timeR, a value from set $\mathbb{N}$, meant to represent the ``time'' at which
the corresponding request (identified by \labelR) is handled by the memory
subsystem. \timeR{} does not have a counterpart in a real implementation, it is
present solely as a specificational device to capture store atomicity. Since a
request corresponding to a response is identified using the field \labelR, for
notational convenience, we will access the fields of the corresponding request
directly from the response. For instance, $r.\addrQ$ for $r \in \Response$
denotes $(r.\labelR).\addrQ$, \etc.

### 2014-02-06 paper/Introduction.tex
A pass through System
Spelling and grammar in Intro + System

Formal verification of cache coherence protocols has traditionally been done
using model-checking based approaches. This signficantly restricts the size as
well as complexity of the system that has to be verified; most model-checking
based approaches can not handle more than one level of hierarchy. To the best
of our knowledge, the largest system that has been verified using
model-checking based approaches has only 12 processors~\cite{Chen:2008:VHC:1559300}.

Apart from scalability concerns, another problem with model-checking based
approaches is that it provides no intuition as to why a protocol is correct, if
the model checker successfully proves a protocol. Developing a new protocol
therefore therefore can not use any intuition that could have been gained from
the first protocol. Moreover the new protocol has to be verified from scratch.

@@

While we have not reached a push-button state in our work, we do develop a
modular proof strategy, where portions of our proof can be used \emph{as is} to
prove other related but distinct protocols. Moreover, the proof gives solid
insights on how the protocol works correctly, which can be used to design other
protocols.

The main contributions of this paper are as follows:
\begin{enumerate}
\item \emph{We give a precise formal definition of a cache coherent system as one
which obeys store atomicity. The definition is abstract in that it does not
rely on any hardware implementation.}
\item \emph{We provide a Coq-based proof of the directory-based MSI cache coherence
protocol with inclusive caches, and no silent drops.}
\item \emph{We present a set of invariants which are not restricted to the MSI
protocol, but would imply store atomicity. Thus, other protocols will also
satisfy store atomicity as long as they obey the invariants.}
\item \emph{We present what we believe to be the key insight behind the
correctness of the MSI protocol. We believe that this property is true also of
other protocols like MESI, MOSI and MOESI, enabling reuse of portions of our
MSI proof for their verification.}
\end{enumerate}
%In practice, verification proceeds in
%two steps. First, a cache coherence protocol is constructed and verified. This
%protocol generally directly represents many of the key concurrent aspects of an
%implementation (\eg{} reads and writes to the system are broken into local
%atomic micro-operations) but ignores implementation details (\eg{} buffer
%sizing and memory access times). Second, once the protocol has been verified,
%it is handed to hardware experts who fill in the missing implementation details
%and realize it at the Register Transfer Level (RTL) of gates, wires, and
%registers, via further breaking down of local micro-operations to make them
%realizable in hardware. This effort is not a one-time translation, as
%performance properties are not easily estimated; a significant amount of
%performance testing is requires to select a good implementation. 
%
%\adamc{I don't understand what this paragraph is saying.}
%From a methodological perspective, this two-step process leaves much to be
%desired. The significant effort of verification is effectively lost by the
%concurrent non-trivial tasks of translation from the protocol representation to
%a hardware description and the necessary refinement to add missing
%implementation details to the protocol.

Protocols, post verification, are often described via Term Rewrite Systems
(TRS) \adamc{citation for TRSes?}, which apply equational logic to describe
systems at an interesting point on the spectrum between declarative and
operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after its
verification, can be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specification can be directly implemented into
efficient hardware. In this way one reduces the correctness of the
implementation to the correctness of the protocol, with no need for separate
reasoning about an implementation. While the current Bluespec compiler does not
support some implementation elements that are common in practice (\eg{}
fairness guarantees, atomic actions running over multiple clock cycles), recent
work has provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a very promising platform for productive implementation of
hardware components, including cache coherent shared memories, starting from a
description meant for making verification easier. Our model of the MSI protocol
in Coq is described as an inductively defined set of transitions which is
directly translatable into Guarded Atomic Actions. In fact, our Coq-based model
is a direct transliteration of the MSI protocol specified in Bluespec.  This
ensures that the final hardware synthesized from the specification is formally
verified.

@@

\noindent\textbf{Paper Outline}. In Section~\ref{Background} we give a
formal definition of cache coherent system as one which obeys store atomicity.
We describe memory subsystems as mathematical functions and define store
atomicity as a property of the input/output pairs of the memory subsystem
function. We describe our MSI protocol and its representation as a
transition relation in Coq in Section~\ref{sec:System}. In
Section~\ref{safety} we discuss the structuring of our proof. We
present the invariants that are sufficient to prove store atomicity and give
the key insight that enables these invariants to be proved.
%We discuss some refinements in
%Section~\ref{Sec:Refinements} and conclude with a general discussion
%of the applicability of our technique.

>>>>

Formal verification of cache coherence protocols has traditionally been done
using model-checking based approaches. This significantly restricts the size as
well as complexity of the system that has to be verified; most model-checking
based approaches cannot handle more than one level of hierarchy. To the best of
our knowledge, the largest system that has been verified using model-checking
based approaches has only 12 processors~\cite{Chen:2008:VHC:1559300}.

Apart from scalability concerns, another problem with model-checking based
approaches is that it provides no intuition as to why a protocol is correct, if
the model checker successfully proves a protocol. Developing a new protocol
therefore cannot use any intuition that could have been gained from the first
protocol. Moreover the new protocol has to be verified from scratch.

@@

While we have not reached a push-button state in our work, we do
develop a modular proof strategy, where portions of our proof can be
used \emph{as is} to prove other related but distinct
protocols. Further, the proof gives solid insights on how the protocol
works correctly, which can be used to design other protocols.

The main contributions of this paper are as follows:
\begin{enumerate}

\item \emph{We give a precise formal definition of a cache coherent
  system as one which obeys store atomicity. The definition is
  abstract in that it does not rely on any hardware implementation.}

\item \emph{We provide a Coq-based proof of the directory-based MSI
  cache coherence protocol with inclusive caches, and no silent
  drops.}

\item \emph{We present a set of invariants which are not restricted to
  the MSI protocol, but would imply store atomicity. Thus, other
  protocols will also satisfy store atomicity as long as they obey the
  invariants.}

\item \emph{We present what we believe to be the key insight behind
  the correctness of the MSI protocol. We believe that this property
  is true also of other protocols like MESI, MOSI and MOESI, enabling
  reuse of portions of our MSI proof for their verification.}

\end{enumerate}

%In practice, verification proceeds in
%two steps. First, a cache coherence protocol is constructed and verified. This
%protocol generally directly represents many of the key concurrent aspects of an
%implementation (\eg{} reads and writes to the system are broken into local
%atomic micro-operations) but ignores implementation details (\eg{} buffer
%sizing and memory access times). Second, once the protocol has been verified,
%it is handed to hardware experts who fill in the missing implementation details
%and realize it at the Register Transfer Level (RTL) of gates, wires, and
%registers, via further breaking down of local micro-operations to make them
%realizable in hardware. This effort is not a one-time translation, as
%performance properties are not easily estimated; a significant amount of
%performance testing is requires to select a good implementation. 
%
%\adamc{I don't understand what this paragraph is saying.}
%From a methodological perspective, this two-step process leaves much to be
%desired. The significant effort of verification is effectively lost by the
%concurrent non-trivial tasks of translation from the protocol representation to
%a hardware description and the necessary refinement to add missing
%implementation details to the protocol.

Protocols, post verification, are often described via Term Rewrite Systems (TRS)
\adamc{citation for TRSes?}, which apply equational logic to describe systems at
an interesting point on the spectrum between declarative and operational
methods.  It was shown by Dave \etal{}~\cite{DNA:CoherenceImplementation} that
such a protocol, after its verification, can be transliterated into
Bluespec~\cite{Bluespec:TFRG}, a hardware description language based on Guarded
Atomic Actions, a restriction of TRS~\cite{Hoe:TCAD}. Bluespec specification can
be directly implemented into efficient hardware. In this way one reduces the
correctness of the implementation to the correctness of the protocol, with no
need for separate reasoning about an implementation. While the current Bluespec
compiler does not support some implementation elements that are common in
practice (\eg{} fairness guarantees, atomic actions running over multiple clock
cycles), recent work has provided key insights toward lifting these
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a promising platform for productive implementation of
hardware, including cache-coherent shared memories, starting from a
verification-empowering description.

Our model of the MSI protocol in Coq is described as an inductively defined set
of transitions which is directly translatable into Guarded Atomic Actions. In
fact, our Coq-based model is a direct transliteration of the MSI protocol
specified in Bluespec.  This ensures that the final hardware synthesized from
the specification is formally verified.

@@

\noindent\textbf{Paper Outline}. In Section~\ref{Background} we give a
formal definition of cache coherent system as one which obeys store atomicity.
We describe memory subsystems as mathematical functions and define store
atomicity as a property of the input/output pairs of the memory subsystem
function. We describe our MSI protocol and its representation as a
transition relation in Coq in Section~\ref{sec:System}. In
Section~\ref{safety} we discuss the structuring of our proof. We
present the invariants that are sufficient to prove store atomicity and give
the key insight that enables these invariants to be proved.

%We discuss some refinements in
%Section~\ref{Sec:Refinements} and conclude with a general discussion
%of the applicability of our technique.

### 2014-02-06 paper/Intuition.tex
Spelling and grammar in Intro + System

We first reduce the proof of store atomicity to 4 invariants, presented in
Section \ref{firstLevel}. These invariants in turn can be reduced to 4 more
invariants presented in Section \ref{secondLevel}. The reason for choosing the 8
 invariants is because they are not specific to MSI protocol
alone; any other protocol which obeys these invariants obeys Theorem
\fullref{storeAtomicity}. We finally present the key insight which we
discovered for proving the second-level invariants in the form of 3 invariants.
We believe that other invalidation-based protocols with inclusive caches (\ie{}
those cache hierarchies where a parent cache contains all the locations that
any of its children contains) like MESI, MOSI and MOESI also
satisfy these 3 fundamental properties, making their verification easier by reusing
portions of our MSI proof.
We informally show why this is the case using the example of MESI,
MOSI and MOESI in Section \ref{mosi}.

>>>>

We first reduce the proof of store atomicity to 4 invariants, presented in
Section~\ref{firstLevel}. These invariants in turn can be reduced to 4 more
invariants presented in Section~\ref{secondLevel}. The reason for choosing the 8
 invariants is because they are not specific to MSI protocol
alone; any other protocol which obeys these invariants obeys Theorem
\fullref{storeAtomicity}. We finally present the key insight which we
discovered for proving the second-level invariants in the form of 3 invariants.
We believe that other invalidation-based protocols with inclusive caches (\ie{}
those cache hierarchies where a parent cache contains all the locations that
any of its children contains) like MESI, MOSI and MOESI also
satisfy these 3 fundamental properties, making their verification easier by reusing
portions of our MSI proof.
We informally show why this is the case using the example of MESI,
MOSI and MOESI in Section~\ref{mosi}.

### 2014-02-06 paper/System.tex
this the the correct version
Spelling and grammar in Intro + System

Each node contains the fields mentioned in Figure \ref{hier}. We will
explain what the fields mean in the following discussion.

\begin{figure}
\includegraphics[scale=.5]{hier}
\caption{A cache hierarchy, and the fields in each node}
\label{hier}
\end{figure}

Input and output happens at the leaves in the form of memory requests
issued from and responses issued to the associated processor. These will be
represented by two mappings, one from $\mathbb{N}$ to requests and the other
from $\mathbb{N}$ to responses. For each leaf $c$, $c.\pos$ represents the
sequence number of the next request to process; it is set to $0$ initially.
\adamc{What is a ``position'' here?  Is ``sequence number'' a better choice?}
\adamc{I hope the diagram explains what these mappings mean, and in general what
the tree structure really means.}

Each node maintains a permission state for each address $a$, $c.\state[a]$, in
addition to the data associated with the address, $c.\data[a]$. The permission
state denotes which operations may be performed on the data. An address may be
in the \emph{read-write} state \Mo{} (for ``modifiable''),
the \emph{read-only} state \Sh{} (for ``shared''), or in the \emph{invalid} state
\In{} (indicating that it can be neither read nor written). The options form a natural
total order of permissions ($\In < \Sh < \Mo$). For a leaf cache, the meaning
of these permission states is straightforward: a load request can be processed
only if the leaf is at least in the \Sh{} state, and a store request can be
processed only if it is in the \Mo{} state. The state of a non-leaf cache indicates
an \emph{upper bound} on the permission level of any cache in the associated subtree.
If a cache does not include a particular location,
we consider that address to have permission \In.  Initially, all addresses in
each cache are in the \In{} state, save for those in the root cache (i.e., main memory),
which are in the \Mo{} state.

@@

In addition to the actual permission states, each cache also maintains a
temporary \emph{wait state} for each address to deal with transitioning between
permission levels. The wait state consists of two parts: (a) a Boolean
representing whether the cache has sent a request to its parent cache earlier
and is waiting for a response (denoted by $c.\wait[a]$ for address $a$) and (b)
one of the three values \Mo, \Sh{} or \In{} representing the state upgrade that
the cache is waiting to make (denoted by $c.\waitS[a]$ for address $a$).
Similarly, each non-leaf cache has the corresponding wait states for its
directory.  $p.\dwait[c][a]$ denotes that cache $p$ is waiting for a response
from its child $c$ for address $a$, and $p.\dwaitS[c][a]$ denotes the state to
which $p$ is waiting for $c$ to downgrade. Both \wait{} and \dwait{} are
+\False{} initially for all caches and addresses,
because no cache is waiting for any response from either its child or its
parent initially.

The behavior of the system is described via the atomic state transitions, which
are effectively a nondeterministic operational semantics. These transitions
have two parts: (a) the guard dictating when the transition can fire, and (b)
the action defining how the state of the system changes when the transition
happens. Logically, these transitions happen one at a time.  Figure~\ref{trans}
gives the atomic transitions for the MIS protocol.  The actions in the
transitions in Figure \ref{trans} show only the specific states that change,
and other state components are assumed to remain the same.  The notation ($x
\Leftarrow y$;) in a transition indicates setting of a specific state $x$ of
the overall system to value $y$.

We first give an intuition on how the transitions work in the common case.
A cache can decide, unsolicitedly \xxx{is this a word?}, to upgrade its state,
in which case it sends an upgrade request to its parent. The parent then makes
a local decision on whether to send a response to the requesting child or not
(based on its directory approximation, and its own state). If its own state is
lower the requested upgrade, then it can not handle the request. The parent
must instead, decide to upgrade its state -- the transition for doing so has
nothing to do with the incoming request.  Once the parent's state is not lower
than the requested upgrade, it makes sure that the rest of its children are
``compatible'' with the requested upgrade. If not, the parent must, send
requests to the uncompatible children to downgrade -- again the transition for
doing so has nothing to do with the incoming request. If the children are
compatible and its own state is higher, then the request can be handled and a
response is sent. Each of the transitions is explained in greater detail below.

Figure~\ref{childside} shows the transitions that result in an upgrade of the
state of a cache. It involves a cache sending an upgrade request to the its
parent if the cache is not already waiting for some response for that address
from the parent (ChildSendReq). This request is received by the parent
(ParentRecvReq). If the directory entries for all of the other children are
\emph{compatible} with the upgrade request, and the parent's permission state has
at least the requested value, it sends an upgrade response to the cache, while
changing the appropriate directory state. The notation \compat{} used in
the ParentRecvReq transition of Figure \ref{trans} is defined as follows:
\begin{multline*}
\compat(p, c, a, s) = \forall i \ne c,
\mylet s_i := p.\dstate[i][a] \myin\\
 (s = \Mo @-> s_i = \In) \wedge (s = \Sh @-> s_i = \Sh) \wedge (s = \In @-> s_i = \Mo)
\end{multline*}
Finally, the child cache receives the response from the parent (ChildRecvResp)
and sets its state while resetting its wait state if the response contains the
desired upgrade.

@@

Because of the presence of voluntary response, a child could have already
downgraded to a state required by a downgrade request from its parent. In such
a case, the child simply drops the downgrade request (ChildDropReq).

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec System Verilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition. Karczmarek \etal~\cite{Karczmarek} has
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.

>>>>

Each node contains the fields mentioned in Figure~\ref{hier}. We will
explain what the fields mean in the following discussion.

\begin{figure}
\centering
\includegraphics[scale=.5]{hier}
\caption{A cache hierarchy, and the fields in each node}
\label{hier}
\end{figure}

Input and output happens at the leaves in the form of memory requests
issued from and responses issued to the associated processor. These will be
represented by two mappings, one from $\mathbb{N}$ to requests and the other
from $\mathbb{N}$ to responses. For each leaf $c$, $c.\pos$ represents the
sequence number of the next request to process; it is set to $0$ initially.
\xxx{Fixed???? \adamc{What is a ``position'' here?  Is ``sequence number'' a better choice?}
\adamc{I hope the diagram explains what these mappings mean, and in general what
the tree structure really means.}}

Each node maintains a permission state for each address $a$, $c.\state[a]$, in
addition to the data associated with the address, $c.\data[a]$. The permission
state denotes which operations may be performed on the data. An address may be
in the \emph{read-write} state \Mo{} (for ``modifiable''),
the \emph{read-only} state \Sh{} (for ``shared''), or in the \emph{invalid} state
\In{} (indicating that it can be neither read nor written). The options form a natural
total order of permissions ($\In < \Sh < \Mo$). For a leaf cache, the meaning
of these permission states is straightforward: a load request can be processed
only if the leaf is at least in the \Sh{} state, and a store request can be
processed only if it is in the \Mo{} state. The state of a non-leaf cache indicates
an \emph{upper bound} on the permission level of any cache in the associated subtree.
If a cache does not include a particular location,
we consider that address to have permission \In.  Initially, all addresses in
each cache are in the \In{} state, save for those in the root cache (\ie{} main memory),
which are in the \Mo{} state.

@@

In addition to the permission states, each cache also maintains a
temporary \emph{wait state} for each address to deal with transitioning between
permission levels. The wait state consists of two parts: (a) a Boolean
representing whether the cache has sent a request to its parent cache earlier
and is waiting for a response (denoted by $c.\wait[a]$ for address $a$) and (b)
one of the three values \Mo, \Sh{} or \In{} representing the state upgrade that
the cache is waiting to make (denoted by $c.\waitS[a]$ for address $a$).
Similarly, each non-leaf cache has the corresponding wait states for its
directory.  $p.\dwait[c][a]$ denotes that cache $p$ is waiting for a response
from its child $c$ for address $a$, and $p.\dwaitS[c][a]$ denotes the state to
which $p$ is waiting for $c$ to downgrade. Both \wait{} and \dwait{} are
+\False{} initially for all caches and addresses,
because no cache is waiting for any response from either its child or its
parent.

The behavior of the system is described via the atomic state transitions, which
are effectively a nondeterministic operational semantics. These transitions
have two parts: (a) the guard dictating when the transition can fire, and (b)
the action defining how the state of the system changes when the transition
happens. Logically, these transitions happen one at a time.  Figure~\ref{trans}
gives the atomic transitions for the MSI protocol.  The actions in the
transitions in Figure~\ref{trans} show only the specific states that change,
and other state components are assumed to remain the same.  The notation ($x
\Leftarrow y$;) in a transition indicates setting of a specific state $x$ of
the overall system to value $y$.

We first give an intuition on how the transitions work in the common case.  A
cache may spontaneously choose to upgrade its state, in which case it sends an
upgrade request to its parent. The parent then makes a local decision on whether
to send a response to the requesting child or not (based on its directory
approximation, and its own state). If its own state is lower the requested
upgrade, then it can not handle the request. The parent must instead, decide to
upgrade its state -- the transition for doing so has nothing to do with the
incoming request.  Once the parent's state is not lower than the requested
upgrade, it makes sure that the rest of its children are ``compatible'' with the
requested upgrade. If not, the parent must, send requests to the incompatible
children to downgrade -- again the transition for doing so has nothing to do
with the incoming request. If the children are compatible and its own state is
higher, then the request can be handled and a response is sent. Each of the
transitions is explained in greater detail below.

Figure~\ref{childside} shows the transitions that result in an upgrade of the
state of a cache. It involves a cache sending an upgrade request to the its
parent if the cache is not already waiting for some response for that address
from the parent (ChildSendReq). This request is received by the parent
(ParentRecvReq). If the directory entries for all of the other children are
\emph{compatible} with the upgrade request, and the parent's permission state has
at least the requested value, it sends an upgrade response to the cache, while
changing the appropriate directory state. The notation \compat{} used in
the ParentRecvReq transition of Figure~\ref{trans} is defined as follows:
\begin{multline*}
\small
\compat(p, c, a, s) = \forall i \ne c,
\mylet s_i := p.\dstate[i][a] \myin\\
 (s = \Mo @-> s_i = \In) \wedge (s = \Sh @-> s_i = \Sh) \wedge (s = \In @-> s_i = \Mo)
\end{multline*}
Finally, the child receives the response from the parent (ChildRecvResp)
and sets its state, resetting its wait state if the response contains the
desired upgrade.

@@

Because of the presence of voluntary response, a child could have already
downgraded to a state required by a downgrade request from its parent. In such
cases, the child simply drops the downgrade request (ChildDropReq).

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec SystemVerilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition. Karczmarek \etal~\cite{Karczmarek} has
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.
